Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-02-10 draft.tex
use abstract from the conference version

	With the ever-increasing threats posed by fake voice synthesis, reliable detection methods
	have become more important than ever. Previous research has tackled this problem using several
	methods, including various features and model architectures, but there remains a need in the
	area of fake voice detection for a method that offers a kind of explainability that can also
	be understood by the average person. We present a method of classification of audio samples as
	synthetic or bona fide using a combination of features with a focus on aspects of a voice
	sample that can be heard by the human ear, perceptible features, that stitches multiple
	subnetworks together in order to evaluate many differently shaped input features in a singular
	model. We then use part of this classifier as a surrogate model to produce explanations using
	the LIME method, with the hope of introducing a kind of explanation into this area that has
	the potential to be understood by a wider range of people and not only people who are already
	familiar with the field. In order to evaluate the contributions of the features to the average
	local explanation, we also introduce two metrics for aggregate evaluations of many local
	explanations, with the goal of making a robust and well justified conclusion about the
	potential usefulness of explanations generated using this method. Experiments were performed
	using the In-the-Wild dataset, a dataset created with a focus on testing generalization of
	previous methods of synthetic audio detection. Several of the surrogate models demonstrated
	strong performance on the dataset, even in comparison to previous methods. However, the
	explanations generated by this method do not seem to offer much potential for useful
	interpretation using the perceptible features, based on an evaluation of aggregated local
	explanations using the proposed metrics, as the imperceptible features have a much larger
	influence on the classification result and only some of the perceptible features had a
	positive influence on the classification correctness. We believe, however, that the proposed
	method offers a promising basis for future research on creating more useful explanations using
	such feature combinations. For the purpose of reproducibility, the source code for these
	experiments is open in the following repository:

>>>>

	With the unprecedented advancement of Generative Artificial Intelligence (GenAI), the threat
	of voice scams using synthetic voices has become a serious concern across various sectors.
	Recent efforts have focused on identifying fake voices through handcrafted features, deep
	learning models, and hybrid approaches. However, most existing methods lack explainability,
	rendering their predictions non-transparent to users. This paper proposes a novel,
	interpretable, and transparent method for fake voice identification by introducing a hybrid
	deep learning model that leverages multiple extracted features. The hybrid model consists of
	two main components: the first component addresses heterogeneous feature spaces by employing
	deep convolutional sub-models tailored to individual features, while the second component, the
	terminus model, utilizes the concatenated representations from the final layers of each
	sub-model as input. The terminus model follows a typical multi-layer perceptron architecture,
	enabling effective integration and classification of the diverse feature representations. To
	enhance interpretability, we decompose the model’s decisions using Local Interpretable
	Model-agnostic Explanations (LIME), exploiting the identical feature representation before the
	concatenation layers to address challenges related to multi-dimensional feature
	representations. To evaluate the importance and trustworthiness of features in the generated
	explanations, we propose two metrics: importance and trust. Extensive experiments are
	conducted on the In-the-Wild dataset, which is designed to test the generalization capability
	of synthetic audio detection methods. The experimental results demonstrate that our approach
	achieves performance comparable to benchmark methods. Furthermore, the results based on our
	proposed metrics conclude that certain perceptible features demonstrate promise for generating
	explanations that are meaningful to general users. For reproducibility, the source code for
	these experiments is available in the following repository:
